Inference in Hidden Markov Models with Explicit State Duration Distributions

### 2011-11-02 tex/introduction.tex
my final (I hope) edits

The explicit state duration HMM (EDHMM) \cite{Rabiner89} was developed to allow explicit parameterization and direct inference of state duration distributions.  Approximate EDHMM estimation and inference can, in some cases, be performed using a modified forward-backward algorithm; specifically if the sequence is short or a tight ``allowable'' duration interval for each state is hard-coded a priori \cite{Yu2006}.   If the sequence is short then forward-backward can be run on a state representation that allows for all possible durations shorter than the observed sequence length.  If the sequence is long then forward-backward only remains computationally tractable if only transitions between durations that lie within pre-specified allowable intervals are considered.   If the true state durations lie outside those intervals then the resulting model estimates will be incorrect: the learned duration distributions will reflect only what is allowed given the pre-specified duration intervals (which may poorly describe the true duration distributions).

Our contribution is the development of a procedure for EDHMM inference that does not require any hard pre-specification of duration intervals and, as a result, is no longer approximate.  The technique we use to do this 
is borrowed from sampling procedures developed for nonparametric  Bayesian HMM variants \cite{vanGael2008}.  Our key insight is simple: the machinery developed for inference in HMMs with a countable number of states is precisely the same as that which is needed for doing inference in an EDHMM with duration distributions over countable support.  So, while the EDHMM is a distinctly parametric model, the tools from nonparametric Bayesian inference can be applied such that black-box inference becomes possible and, in fact, efficient.

>>>>

The explicit state duration HMM (EDHMM) \cite{Rabiner89} was developed to allow explicit parameterization and direct inference of state duration distributions.   EDHMM estimation and inference can be performed using the forward-backward algorithm; specifically if the sequence is short or a tight ``allowable'' duration interval for each state is hard-coded a priori \cite{Yu2006}.   If the sequence is short then forward-backward can be run on a state representation that allows for all possible durations up to the observed sequence length.  If the sequence is long then forward-backward only remains computationally tractable if only transitions between durations that lie within pre-specified allowable intervals are considered.   If the true state durations lie outside those intervals then the resulting model estimates will be incorrect: the learned duration distributions will reflect only what is allowed given the pre-specified duration intervals. % (which may poorly describe the true duration distributions).

Our contribution is the development of a procedure for EDHMM inference that does not require any hard pre-specification of duration intervals and, as a result, is no longer approximate.  The technique we use to do this 
is borrowed from sampling procedures developed for nonparametric  Bayesian HMM variants \cite{vanGael2008}.  Our key insight is simple: the machinery developed for inference in HMMs with a countable number of states is precisely the same as that which is needed for doing inference in an EDHMM with duration distributions over countable support.  So, while the EDHMM is a distinctly parametric model, the tools from nonparametric Bayesian inference can be applied such that black-box inference becomes possible and, in practice, efficient.
